Abstractions for Multi-Sorted Substitutions

### 2024-03-13 paper/paper.tex
Update.

  Given a \ACode{\ACon{CTraversal}}, we can prove the lemmas about interactions
  of multiple maps, which are required for substitution preserves typing:
  \begin{itemize}
  \item 
    A map \ACode{ϕ} followed by a weakening is equivalent to a
    weakening followed by \ACode{ϕ} that has been lifted over the weakened
    variable:
    \LibCode\KCommLiftWeakenTraverse

@@

  \AppCode\FFusion
  \AppCode\FFusionProofInteresting
  The interesting cases are:
  \begin{itemize}
  \item
    variables, where we need to use the \ACode{\AField{\&/⋯-⋯}} lemma
    provided by the \ACode{\ACon{CKit}}.
  \item
    bindings, where the traversal operation \ACode{\AField{\_⋯\_}}
    needs to lift the map via \ACode{\AField{\_↑\_}}, requiring us to
    distribute the lifting over the composition using
    \ACode{\AField{dist-↑-·}}.
  \end{itemize}

@@

  In the context of multi-sorted syntax, the notion of a type can be
  described as a mapping between sorts.
  For System F, the expression sort \ACode{\ACon{𝕖}} maps to the type sort
  \ACode{\ACon{𝕥}} and the sort \ACode{\ACon{𝕥}} maps to the kind sort
  \ACode{\ACon{𝕜}}.
  The following structure is used to teach our framework about types:
  \LibCode\KTypeSorts
  For System F the instantiation is
  \AppCode\FTypes
  There are two things to discuss:
  \begin{enumerate}
  \item
    The \ACode{\AField{↑ᵗ}} function maps a sort of arbitrary sort type,
    to a sort of a potentially different sort type, which is expressed
    by the use of an existential.
    For System F we require this generality, as the sort
    \ACode{\ACon{𝕥}} can have variables, whereas its corresponding type
    sort \ACode{\ACon{𝕜}} cannot.
  \item
    Some sorts, like \ACode{\ACon{𝕜}}, do not have corresponding type
    sorts, but we still need to provide one, as \ACode{\AField{↑ᵗ}} is a
    total function.
    For such sorts, we can simply use arbitrary sort types, as the
    formalization will never use them.
  \end{enumerate}
  To hide the existential, we define \ACode{S \ACon{∶⊢} s}, which represents
  the type for a term \ACode{S \ACon{⊢} s}.
  \LibCode\KTypes

@@

  As we want our framework to support dependent types, we allow a type returned
  by the typing context to use all variables bound previously in the typing context.
  This is achieved by the \ACode{drop-∈} function, which cuts of all
  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.

>>>>

  \noindent Given a \ACode{\ACon{CTraversal}}, we can prove the usual lemmas about interactions
  of multiple maps:
  \begin{itemize}
  \item 
    A map \ACode{ϕ} followed by a weakening is equivalent to a
    weakening followed by \ACode{ϕ} that has been lifted over the weakened
    variable:
    \LibCode\KCommLiftWeakenTraverse

    % A map \ACode{ϕ} followed by a weakening is equivalent to a
    % weakening followed by \ACode{ϕ} that has been lifted over the weakened
    % variable:
    % \LibCode\KCommLiftWeaken
    % % \LibCode\KCommLiftWeakenProof
    % By applying \ACode{\AField{fusion}} to this lemma, we get the same
    % statement for multiple map applications:
    % \LibCode\KCommLiftWeakenTraverse
    % % \LibCode\KCommLiftWeakenTraverseProof
  \item 
    A weakening followed by a singleton substitution act as an identity map:
    \LibCode\KWeakenCancelsSingleTraverse

@@

  \AppCode\FFusion
  \AppCode\FFusionProofInteresting
  We only show the interesting cases, which are:
  \begin{itemize}
  \item
    variables, where we need to use the \ACode{\AField{\&/⋯-⋯}} lemma
    provided by the \ACode{\ACon{CKit}}; and
  \item
    bindings, where the traversal operation \ACode{\AField{\_⋯\_}}
    needs to lift the map via \ACode{\AField{\_↑\_}}, requiring us to
    distribute the lifting over the composition using
    \ACode{\AField{dist-↑-·}}.
  \end{itemize}

@@

  In the context of multi-sorted syntax, the notion of a type can be
  described as a mapping between sorts.
  For System F, the expression sort \ACode{\ACon{𝕖}} maps to the type sort
  \ACode{\ACon{𝕥}}, and the sort \ACode{\ACon{𝕥}} maps to the kind sort
  \ACode{\ACon{𝕜}}.
  The following structure is used to teach our framework about types:
  \LibCode\KTypeSorts
  For System F the instantiation is
  \AppCode\FTypes
  There are two things to discuss:
  \begin{enumerate}
  \item
    The \ACode{\AField{↑ᵗ}} function maps a sort of arbitrary sort type,
    to a sort of a potentially different sort type, which is expressed
    by the use of an existential.
    For System F we require this generality, as the sort
    \ACode{\ACon{𝕥}} can have variables, whereas its corresponding type
    sort \ACode{\ACon{𝕜}} cannot.
  \item
    Some sorts, like \ACode{\ACon{𝕜}}, do not have corresponding type
    sorts, but we still need to provide one, as \ACode{\AField{↑ᵗ}} is a
    total function.
    For such sorts, we can simply use arbitrary sort types, as the
    formalization will never use them.
  \end{enumerate}
  To hide the existential, we define \ACode{S \ACon{∶⊢} s}, which represents
  the type for a term \ACode{S \ACon{⊢} s}.
  \LibCode\KTypes

@@

  As we want our framework to support dependent types, we allow a type returned
  by the typing context to use all variables bound previously in the typing context.
  This is achieved by the \ACode{drop-∈} function, which cuts off all
  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
